---
title: "Homework 1"
date: July 5 2023
author: Jakey Lebwohl
output: github_document
---


# Problem 1

Before getting started, we will need the _tidyverse_ library.

```{r message = FALSE}
library(tidyverse)
```

The following code chunk loads the penguins dataset.

```{r}
data("penguins", package = "palmerpenguins")
```

This dataset has `r ncol(penguins)` variables: for each of the `r nrow(penguins)` penguins, the dataset includes the species, island, bill length (in mm), bill depth (mm), flipper length (mm), body mass (g), sex, and year.  
  
In order to find the mean flipper length of the penguins in this data set, we can use the _mean_ function, excluding penguins without flippers: 
```{r}
flipper_mean = mean(penguins$flipper_length_mm, na.rm=TRUE)
round(flipper_mean, digits = 2)
```

We can make a scatterplot too. The following code chunk creates a scatterplot with flipper length (mm) as the y variable and bill length (mm) as the x variable.

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() # make scatterplot
ggsave("flippers_vs_bills.pdf") # save scatterplot as pdf
```

# Problem 2

We will now create a random sample of size 10 from a normal distribution and assign certain variables to each number.

```{r}
sample_df = tibble(
  value = rnorm(10),
  value_pos = value > 0,
  value_word = c("2", "3", "5", "7", "11", "13", "17", "19", "23", "29"),
  value_factor = factor(c("one", "prime", "prime", "composite", "prime", "composite", "prime", "composite", "composite", "composite"))
)
```

```{r include = FALSE}
is_isnt = c("**is**", "**isn't**")
```

We _can_ take the mean of the set of 10 numerical values: the mean is `r round(mean(pull(sample_df, value)), digits = 3)`. This mean should be close to 0: there is a 50% probability the mean will be between -0.213 and 0.213. In this case, it `r is_isnt[ceiling(abs(mean(pull(sample_df, value))) - 0.213 + 1)]`.

We _can_ take the mean of the set of logical values. This is because TRUE and FALSE are interchanged with 0 and 1:

```{r}
pull(sample_df, value_pos)
mean(pull(sample_df, value_pos))
```

We _cannot_ take the mean of the set of characters:
```{r}
mean(pull(sample_df, value_word))
```

We _cannot_ take the mean of the set of factors:
```{r}
mean(pull(sample_df, value_factor))
```

Using _as.numeric_, we can convert many types of variables to numeric values:
```{r, message==FALSE}
x = as.numeric(pull(sample_df, value_pos))
y = as.numeric(pull(sample_df, value_word))
z = as.numeric(pull(sample_df, value_factor))
```

When this function is applied to a logical value, TRUE gets converted to 1 and FALSE gets converted to 0.  
  
When this function is applied to a character value, any numbers in the character value get converted to numerical values. Every other character value is converted to NA.  
  
When this function is applied to a factor, it gives each factor a different natural number (apparently dependent on its position among all other factors alphabetically).